Modular transmission line probes for microfluidic nuclear magnetic resonance spectroscopy and imaging

### 2019-01-21 ms5n17-tlp-mn-180216-001.tex
updated abstract; cleanup

\begin{abstract}
Microfluidic NMR (Nuclear Magnetic Resonance) can probe chemical or bio-chemical reactions, including the metabolism of cells, and tissues non-invasively in a tightly controlled and tailored microfluidic environment. We present a dual-channel modular probe assembly for  high efficiency microfluidic-NMR. The samples are contained in exchangeable, flow compatible planar microfluidic devices. The detector including the tuning and matching circuitry of the probe is modular and allows the use of different detectors optimised for different experimental conditions on the same probe base. The complete probe can be built from easily available parts. The probe body is made from prefabricated aluminium profiles and few precision machined parts, while the probe circuit and detector are made from PCB (printed circuit board) materials. The probe has a limit of detection of 1.4 nmol s$^{-1/2}$, resolution of 3.35 Hz and, an excellent B$_{1}$ homogeneity. The probe is optimised for proton detection and can be tuned and matched simultaneously to proton and X channel (\textsuperscript{13}C or \textsuperscript{15}N) frequencies. We have successfully acquired \textsuperscript{1}H-\textsuperscript{13}C and \textsuperscript{1}H-\textsuperscript{15}N heteronuclear correlation spectrum (HSQC). Particularly, \textsuperscript{1}H-\textsuperscript{15}N HSQC spectrum of 1 mM \textsuperscript{15}N labeled ubiquitin from 2.5 $\mu$l of sample volume.
\end{abstract}
\maketitle

@@

NMR (Nuclear Magnetic Resonance) is a well established technique used in chemical 
analysis~\cite{NMR-chemical-1995, hills1994magnetic, rabenstein1991quantitative},
molecular structure determination~\cite{wuthrich1990protein}, metabolomics~\cite{nmr-metabolomics-future-2017,nmr-metabolomics-2016}, reaction 
monitoring~\cite{maiwald2004quantitative} etc. Microfluidic lab-on-a-chip (LOC) devices 
are increasingly finding applications in chemistry and the life 
sciences~\cite{whitesides2006origins,mark2010microfluidic}. 
They provide a convenient way to integrate complex chemical and biochemical processes. 
In particular, microfluidic technology allows detailed control over growth conditions 
in the culture of cells, cell aggregates, and small organisms~\cite{cellonchip-2006,cellonchip-review,tissueonchip-2008}. 
Lab-on-a-chip devices can also enable high experimental 
throughput by parallelisation. 
In spite of its analytical power, NMR has so far only rarely been used in microfluidic 
systems. In general, information is extracted from microfluidic devices by spectroscopic 
techniques at the end point of the experiment. Typically, 
this is done using fluorescent probes, which bind to one or a few analytes of interest. 

@@

femtomolar concentrations. However, it is usually destructive, since the sample is 
affected by the addition of the marker, and can only be done once, at the end of the 
experiment. Following kinetic processes therefore requires using multiple samples. By 
contrast, NMR spectroscopy is non-invasive, generic (virtually all organic 
molecules give an NMR signal), and highly specific (no two metabolites share 
the same NMR spectrum).

NMR spectroscopy can provide a useful complementary readout capability 
for microfluidic devices. For example, the metabolic activity of cell cultures 
can be followed non-invasively by NMR throughout the course of the experiment~\cite{cellnmr-2015}, 
while effects on cell differentiation can still be analysed by a (destructive) 
transcriptomic analysis at the end. Unfortunately, the sensitivity 
of NMR is much lower than that of optical techniques. 
The small volumes involved in microfluidic systems
 (typically in the range of 1~$\mu$l) 
compound this problem. As is well known, the mass limit of detection of 
inductive NMR detectors improves as they are scaled down~\cite{Olson1995}. 
On this basis, a number of different micro-NMR detector designs have been 
proposed and characterised in the literature~\cite{utz2012review,micronmr2014review}. 
In most cases, these are based on micro-solenoid~\cite{SUBRAMANIAN1998,Pines2007}, 
planar~\cite{Maguire2007,dieter2008,EHRMANN200}, or stripline 
detectors~\cite{stripline_jan}. 
Significant research has gone into developing micro-NMR detectors for flow probes, 
where the sample container takes the form of a fixed capillary into which the sample
must be injected. Only a few designs that allow insertion of a removable microfluidic 
device into the detector have been described~\cite{Spengler-2014,Spengler-2016,gream_2016}.

In this paper we present a novel modular probe designed for generic 
microfluidic-NMR experiments. The design is based on a transmission line 
detector~\cite{stripline_jan,gream_2016} using inexpensive 
printed circuit board (PCB) technology. A modular design ensures that detectors of 
different geometry and size can be easily exchanged. 
The NMR detector (including the tuning and matching circuit) is connected to the probe 
skeleton through detachable connectors. Modularity of the probe not only allows 
fast optimisation of the detector but also facilitates the use of dedicated detectors 
for different microfluidic devices on a single probe base. 
Once the probe base is in place, additional detectors can be manufactured 
within 2-3 days for less than $\sim$200 $\geneuro$. 
The probe is designed for proton detected double resonance. 
The second channel can be tuned to different nuclei such as $^{13}$C, $^{15}$N, or 
$^{31}$P. The probe skeleton is made almost entirely from easily obtainable raw materials 
such as aluminium profiles of standard size, with only a small number of 
custom-machined parts holding the structure together. 
Detailed construction drawings for the probe are provided in the supporting 
information, and a complete set of CAD files is available for download at \todo{needs to be discussed}.
The planar geometry of the detector accommodates exchangeable microfluidic devices 
which can be designed freely for specific experimental requirements as long as the 
overall shape of the device is maintained.

@@

Microfluidic devices are made from thermally bonding PMMA sheets.
Briefly, the devices consist of a 0.5~mm thick middle layer between a top and bottom layer of 
0.2~mm thickness each. The sample chamber and fluidic network are cut into the middle layer.
The fluidic network on the device can be designed freely as long as the overall shape of the
device is preserved, and the location of the sample chamber is maintained.

@@

\begin{figure}
\centering
\includegraphics[width=.7\linewidth,keepaspectratio=true]{./figures/pcb-fieldmap-colorbar.png} 
\caption{B\textsubscript{0} field map of the sample chamber for FR4 (A) and Rogers RO3035 (B) PCB material. Both the field maps were constructed from 2 gradient echo images acquired on a chip filled with 130 mM sodium acetate dissolved in H\textsubscript{2}O through same parameters and pulse sequence.}
\label{fig:field map} 
\end{figure}

@@

@@

\begin{figure}
\centering
\includegraphics[width=.7\linewidth,keepaspectratio=true]{./figures/media-180125-005-bigger-font.png} 
\caption{Spectrum of culture growth media recorded in 256 scans with water presaturation. The total experimental time was 20 min for 2.5 $\mu$l sample volume.}
\label{fig:media-spec} 
\end{figure}

\begin{figure}
\centering
\includegraphics[width=.7\linewidth,keepaspectratio=true]{./figures/ms5n17-tlp-sp-180612-HSQCspect.png} 
\caption{ $^1$H-$^{13}$C HSQC spectrum of $^{13}$C labeled glucose (left) and $^1$H-$^{15}$N HSQC of $^{15}$N labeled Ubiquitin (right) both acquired from 2.5$\mu$l sample volume at 14.1 T. Glucose HSQC spectrum was acquired in 12 minutes from a 100 mM sample. Ubiquitin HSQC spectrum was acquired in 400 minutes from a 1 mM sample.}
\label{fig:HSQC} 
\end{figure}
%$ 
The probe can also be used to acquire micro magnetic resonance (MR) images using a Bruker micro-gradient unit. 
To this effect, the aluminium probe sleeve must be replaced by an electrically insulating material in order to 
avoid image distortions due to eddy currents in the sleeve. 
\fig{fig:tisli} shows MR images of mouse liver tissue slice 
immersed in DMEM cell growth medium.
The slice was placed in a microfluidic device with circular sample chamber centred at the detector location, as 
shown in \fig{fig:tisli}(A). 
The image shown in \fig{fig:tisli}(B) was acquired using a FLASH pulse sequence with
$t_R=120$~ms and $t_E=15.3$~ms. The slice thickness of  1~mm encompasses the entire thickness of the tissue; the 
lateral dimensions of all MR images in \fig{fig:tisli} are 512$\times$512 pixels
at a resolution of 30~$\mu$m.
\fig{fig:tisli}(C) shows a RARE image using the same parameters, with 8 echos
acquired per scan, thus reducing the acquisition time to 8 minutes.
A spin echo image is shown in \fig{fig:tisli}(D), acquired over 25~min. Finally,
\fig{fig:tisli}(E) shows a $B_0$ field map, which has been obtained by acquiring
two separate FLASH images with echo times $t_E=5.3$~ms and 15.3~ms, respectively.
, corresponding to local field variations of $\Delta B/B_0=...$ 
\todo{compute value and insert here}.

\begin{figure}
\centering
\includegraphics[width=.8\linewidth,keepaspectratio=true]{./figures/ms5n17-tisli-im-180511.png} 
\caption{MR images of a freshly cut live slice from a mouse. A: optical micrograph; B: Gradient echo image acquired in 1 minute; C: RARE8 image acquired in 8 minutes; D: Spin echo image acquired in 30 minutes; E: magnetic field map.}
\label{fig:tisli} 
\end{figure}

>>>>

\begin{abstract} Microfluidic NMR spectroscopy can probe chemical and
bio-chemical processes non-invasively in a tightly controlled environment. We
present a dual-channel modular probe assembly for  high efficiency microfluidic
NMR spectroscopy and imaging. It is compatible with a wide range of microfluidic
devices,  without constraining the fluidic design. It collects NMR signals from
a designated sample volume on the device with high sensitivity and resolution.
Modular design allows adapting the detector geometry to different experimental
conditions with minimal cost, by using the same probe base. The complete probe
can be built from easily available parts. The probe body mainly consists of
prefabricated aluminium profiles, while the probe circuit and detector are made
from printed circuit boards. We demonstrate a double resonance HX probe with a
limit of detection of 1.4 nmol s$^{-1/2}$ for protons at 600~MHz, resolution of
3.35 Hz, and excellent B$_{1}$ homogeneity. We have successfully acquired
\textsuperscript{1}H-\textsuperscript{13}C and
\textsuperscript{1}H-\textsuperscript{15}N heteronuclear correlation spectra
(HSQC), including a \textsuperscript{1}H-\textsuperscript{15}N HSQC spectrum of
1 mM \textsuperscript{15}N labeled ubiquitin in 2.5 $\mu$l of sample volume.
\end{abstract} \maketitle

@@

NMR (Nuclear Magnetic Resonance) is a well established technique used in
chemical analysis~\cite{NMR-chemical-1995, hills1994magnetic,
rabenstein1991quantitative}, molecular structure
determination~\cite{wuthrich1990protein},
metabolomics~\cite{nmr-metabolomics-future-2017,nmr-metabolomics-2016}, reaction
monitoring~\cite{maiwald2004quantitative} etc. Microfluidic lab-on-a-chip (LOC)
devices are increasingly finding applications in chemistry and the life
sciences~\cite{whitesides2006origins,mark2010microfluidic}. They provide a
convenient way to integrate complex chemical and biochemical processes. In
particular, microfluidic technology allows detailed control over growth
conditions in the culture of cells, cell aggregates, and small
organisms~\cite{cellonchip-2006,cellonchip-review,tissueonchip-2008}.
Lab-on-a-chip devices can also enable high experimental throughput by
parallelisation. In spite of its analytical power, NMR has so far only rarely
been used in microfluidic systems. In general, information is extracted from
microfluidic devices only at the end point of the experiment. Typically, this is
done using fluorescent probes, which bind to one or a few analytes of interest.

@@

femtomolar concentrations. However, it is usually destructive, since the sample
is affected by the addition of the marker, and can only be done once, at the end
of the experiment. Following kinetic processes therefore requires using multiple
samples. By contrast, NMR spectroscopy is non-invasive, generic (virtually all
organic molecules give an NMR signal), and highly specific (no two metabolites
share the same NMR spectrum), and can therefore provide a useful complementary
readout capability for microfluidic devices. For example, the metabolic activity
of cell cultures can be followed non-invasively by NMR throughout the course of
the experiment~\cite{cellnmr-2015}, while effects on cell differentiation can
still be analysed by a (destructive) transcriptomic analysis at the end.
Unfortunately, the sensitivity of NMR is much lower than that of optical
techniques. The small volumes involved in microfluidic systems (typically in the
range of 1~$\mu$l) compound this problem. As is well known, the mass limit of
detection of inductive NMR detectors improves as they are scaled
down~\cite{Olson1995}. On this basis, a number of different micro-NMR detector
designs have been proposed and characterised in the
literature~\cite{utz2012review,micronmr2014review}. In most cases, these are
based on micro-solenoid~\cite{SUBRAMANIAN1998,Pines2007},
planar~\cite{Maguire2007,dieter2008,EHRMANN200}, or stripline
detectors~\cite{stripline_jan}. Significant research has gone into developing
micro-NMR detectors for flow probes, where the sample container takes the form
of a fixed capillary into which the sample must be injected. Only a few designs
that allow insertion of a removable microfluidic device into the detector have
been described~\cite{Spengler-2014,Spengler-2016,gream_2016}.

In this paper we present a novel modular probe designed for generic
microfluidic-NMR experiments. The design is based on a transmission line
detector~\cite{stripline_jan,gream_2016} using inexpensive
printed circuit board (PCB) technology. A modular design ensures that detectors of
different geometry and size can be easily exchanged.
The NMR detector (including the tuning and matching circuit) is connected to the probe
skeleton through detachable connectors. Modularity of the probe not only allows
fast optimisation of the detector but also facilitates the use of dedicated detectors
for different microfluidic devices on a single probe base.
Once the probe base is in place, additional detectors can be manufactured
within 2-3 days for less than $\sim$200 $\geneuro$.
The probe is designed for proton detected double resonance.
The second channel can be tuned to different nuclei such as $^{13}$C, $^{15}$N, or
$^{31}$P. The probe skeleton is made almost entirely from easily obtainable raw materials
such as aluminium profiles of standard size, with only a small number of
custom-machined parts holding the structure together.
Detailed construction drawings and a complete set of CAD files is
available for download at \cite{github-probe-download-2019}.
The planar geometry of the detector accommodates exchangeable microfluidic devices
which can be designed freely for specific experimental requirements as long as the
overall shape of the device is maintained.

@@

Microfluidic devices are made from inexpensive PMMA sheets by laser cutting and
thermal bonding, as described in detail elsewhere \cite{yilmaz_bonding}.
Briefly, the devices consist of a 0.5~mm thick middle layer between a  top and
bottom layer of 0.2~mm thickness each. The sample chamber and fluidic network
are cut into the middle layer. The fluidic network on the device can be designed
freely as long as the overall shape of the device is preserved, and the location
of the sample chamber is maintained.

@@

\begin{figure} \centering
\includegraphics[width=.9\linewidth,keepaspectratio=true]{./figures/pcb-fieldmap-colorbar.png}
\caption{B\textsubscript{0} field map of the sample chamber for FR4 (A) and
Rogers RO3035 (B) PCB material. Both the field maps were constructed from 2
gradient echo images acquired on a chip filled with 130 mM sodium acetate
dissolved in H\textsubscript{2}O through same parameters and pulse sequence.}
\label{fig:field map}
\end{figure}

@@

\begin{figure*}
\centering
\includegraphics[width=.7\linewidth,keepaspectratio=true]{./figures/media-180125-005-bigger-font.png}
\caption{Spectrum of culture growth media recorded in 256 scans with water presaturation. The total experimental time was 20 min for 2.5 $\mu$l sample volume.}
\label{fig:media-spec}
\end{figure*}

@@

\begin{figure} \centering
\includegraphics[width=\linewidth,keepaspectratio=true]{./figures/ms5n17-tlp-sp-180612-HSQCspect.png}
\caption{ $^1$H-$^{13}$C HSQC spectrum of $^{13}$C labeled glucose (left) and
$^1$H-$^{15}$N HSQC of $^{15}$N labeled Ubiquitin (right) both acquired from
2.5$\mu$l sample volume at 14.1 T. Glucose HSQC spectrum was acquired in 12
minutes from a 100 mM sample. Ubiquitin HSQC spectrum was acquired in 400
minutes from a 1 mM sample.} \label{fig:HSQC} \end{figure}
%$
The probe can also be used to acquire micro magnetic resonance (MR) images using
a Bruker micro-gradient unit. To this effect, the aluminium probe sleeve must be
replaced by an electrically insulating material in order to avoid image
distortions due to eddy currents in the sleeve. \fig{fig:tisli} shows MR images
of mouse liver tissue slice immersed in DMEM cell growth medium. The slice was
placed in a microfluidic device with circular sample chamber centred at the
detector location, as shown in \fig{fig:tisli}(A). The image shown in
\fig{fig:tisli}(B) was acquired using a FLASH pulse sequence with $t_R=120$~ms
and $t_E=15.3$~ms. The slice thickness of  1~mm encompasses the entire thickness
of the tissue; the lateral dimensions of all MR images in \fig{fig:tisli} are
512$\times$512 pixels at a resolution of 30~$\mu$m. \fig{fig:tisli}(C) shows a
RARE image using the same parameters, with 8 echos acquired per scan, thus
reducing the acquisition time to 8 minutes. A spin echo image is shown in
\fig{fig:tisli}(D), acquired over 25~min. Finally, \fig{fig:tisli}(E) shows a
$B_0$ field map, which has been obtained by acquiring two separate FLASH images
with echo times $t_E=5.3$~ms and 15.3~ms, respectively. , corresponding to local
field variations of $\Delta B/B_0=...$ \todo{compute value and insert here}.

\begin{figure}
\centering
\includegraphics[width=\linewidth,keepaspectratio=true]{./figures/ms5n17-tisli-im-180511.png}
\caption{MR images of a freshly cut live slice from a mouse. A: optical micrograph; B: Gradient echo image acquired in 1 minute; C: RARE8 image acquired in 8 minutes; D: Spin echo image acquired in 30 minutes; E: magnetic field map.}
\label{fig:tisli}
\end{figure}
